---
title: "Regions of Interest"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Regions of Interest}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = T, comment = "#>")
library(neuroim)
```


Regions of interest
===================

## Creating a Spherical ROI

In **neuroim** there is basic support for creating regions of interest (ROI). To create a spherical ROI around a central point, we need an existing object of type `BrainVolume` or `BrainSpace`.

To create a spherical region of interest with a 5mm radius around a central voxel at i=20, j=20, k=20, we can do the following:

```{r}
      # attach MNI BrainSpace instance
      
      data("MNI_SPACE_1MM")
      
      # we create a spherical ROI centered around voxel coordinates [20,20,20] with a 5mm radius, 
      # filling all values in the ROI with 100.

      sphere <- RegionSphere(MNI_SPACE_1MM, c(20,20,20), radius=5, fill=100)
      
      # to extract the voxel coordinates of the sphere:
      
      vox <- coords(sphere)
      
      # to get the values at the coordinate locations
      
      vals <- values(sphere)
      all.equal(vals, rep(100, length(vals)))   
```
      
## Creating a Spherical ROI around an MNI coordinate

To create a spherical ROI centered around an real coordinate in mm, we need to first convert the real-valued coordinate to a voxel-based coordinate.
Suppose our real-world coordinate is at -50, -28, 10 in MNI space.

```{r}

    
    
    rpoint <- c(-50,-28,10)
    
    # Because RegionSphere takes a coordinate in voxel units, 
    # we need to convert to the real-world MNI coordinate to grid coordinates.
    
    vox <- coordToGrid(MNI_SPACE_1MM, rpoint)
    sphere <- RegionSphere(MNI_SPACE_1MM, vox, radius=10, fill=1)
    dim(coords(sphere))
    
    # convert back to MNI coordinates
    
    mnicoords <- indexToCoord(MNI_SPACE_1MM, indices(sphere))
    
    ## compute center of mass of MNI coords in ROI (should be close to original coordinate)
    centerOfMass <- colMeans(mnicoords)
    centerOfMass
    
```
    
## Converting an ROI to a SparseBrainVolume

We may want to convert a region of interest to a BrainVolume instance. But we don't want to store every value in dense grid. Here we can make use of the `SparseBrainVolume` class which only stores non-zero values. 

```{r}
    
    sphere <- RegionSphere(MNI_SPACE_1MM, c(50,50,50), radius=10, fill=1)
    sparsevol <- SparseBrainVolume(values(sphere),MNI_SPACE_1MM,indices=indices(sphere))
    
    sum(sparsevol) == sum(values(sphere))
    
    all(dim(sparsevol) == dim(MNI_SPACE_1MM))
    
    
    
```






    
